A sheaf--function dictionary for algebraic tori over local fields

### 2013-08-18 shanghai.tex
Adding proof of corollary and some references

\begin{corollary}
 If $G$ is a commutative, \'etale algebraic group over $\Fq$ and $G(\bFq)$ is finitely
 generated then
  \[
  \trFrob{G} : \QCiso{G} \to G(\Fq)^*
  \]
  is an isomorphism of groups.
\end{corollary}

@@

\begin{proof}
  Recall the short exact sequence \eqref{eq:pi0} in the category of smooth group
  schemes, locally of finite type over $\Fq$, defining the component
  group scheme for $G$:
  \[
  \begin{tikzcd}
    1 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
  \end{tikzcd}
  \]
  Let
  \[
  \begin{tikzcd}[row sep=30]
    {}& \ker \trFrob{\pi_0(G)} \dar & \arrow[dashed]{d} \ker \trFrob{G} & \ker \trFrob{G^0} \dar & \\
    0 \rar & \QCiso{\pi_0(G)} \rar \dar{\trFrob{\pi_0(G)}}
    & \QCiso{G} \rar \dar{\trFrob{G}} & \QCiso{G^0} \rar \dar{\trFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar \dar
    & \arrow[dashed]{d} G(\Fq)^* \rar & G^0(\Fq)^* \rar \dar & 0\\
    & \coker \trFrob{\pi_0(G)} & \coker \trFrob{G} &  \coker \trFrob{G^0} &
  \end{tikzcd}
  \]
  be the commutative diagram of abelian groups obtained by applying
  Proposition~\ref{prop:pullback} to \eqref{eq:pi0}.  Since $G^0$ is a connected algebraic group
  \cite{vdGeer-Moonen:AbelianVarieties}*{3.17},
  Proposition~\ref{prop:connected} implies that $\ker \trFrob{G^0} =0$ and $\coker \trFrob{G^0}=0$.
  By Propositions~\ref{prop:etale} and \ref{prop:X}, $\ker \trFrob{\pi_0(G)}=0$
  and $\coker \trFrob{\pi_0(G)}=0$.

>>>>

\begin{corollary}
 If $G$ is a commutative, \'etale group scheme over $\Fq$ and $G(\bFq)$ is finitely
 generated then
  \[
  \trFrob{G} : \QCiso{G} \to G(\Fq)^*
  \]
  is an isomorphism of groups.
\end{corollary}
\begin{proof}
This follows from Proposition~\ref{prop:etale} by setting $X=G(\bFq)$ in Proposition~\ref{prop:X}.
\end{proof}

@@

\begin{proof}
  Recall the short exact sequence \eqref{eq:pi0} in the category of smooth group
  schemes, locally of finite type over $\Fq$, defining the component
  group scheme for $G$:
  \[
  \begin{tikzcd}
    1 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
  \end{tikzcd}
  \]
  Let
  \[
  \begin{tikzcd}[row sep=30]
    {}& \ker \trFrob{\pi_0(G)} \dar & \arrow[dashed]{d} \ker \trFrob{G} & \ker \trFrob{G^0} \dar & \\
    0 \rar & \QCiso{\pi_0(G)} \rar \dar{\trFrob{\pi_0(G)}}
    & \QCiso{G} \rar \dar{\trFrob{G}} & \QCiso{G^0} \rar \dar{\trFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar \dar
    & \arrow[dashed]{d} G(\Fq)^* \rar & G^0(\Fq)^* \rar \dar & 0\\
    & \coker \trFrob{\pi_0(G)} & \coker \trFrob{G} &  \coker \trFrob{G^0} &
  \end{tikzcd}
  \]
  be the commutative diagram of abelian groups obtained by applying
  Proposition~\ref{prop:pullback} to \eqref{eq:pi0}.  Since $k$ is perfect
  \cite{stacks-project}*{\href{http://stacks.math.columbia.edu/tag/035X}{Tag 035X}} implies that
  \David{Need another reference that an algebraic group over an algebraically closed field is reduced}
  $G$ is reduced.  Since $G^0$ is a finite type
  \cite{vdGeer-Moonen:AbelianVarieties}*{3.17}, it is thus a connected algebraic group and 
  Proposition~\ref{prop:connected} implies that $\ker \trFrob{G^0} =0$ and $\coker \trFrob{G^0}=0$.
  By Propositions~\ref{prop:etale} and \ref{prop:X}, $\ker \trFrob{\pi_0(G)}=0$
  and $\coker \trFrob{\pi_0(G)}=0$.
